\documentclass[12pt,a4paper]{paper}
\usepackage[utf8]{inputenc}
\usepackage[spanish]{babel}
\usepackage{amsmath}
\usepackage{enumitem}
\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage[left=1cm,right=1cm,top=1cm,bottom=2cm]{geometry}
\usepackage{Sweave}
\begin{document}
\title{STAT636 - Homework 1\\\small{Daniel Osorio - dcosorioh@tamu.edu\\Department of Veterinary Integrative Biosciences\\Texas A\&M University}}
\maketitle
\input{HW1_DanielOsorio-concordance}
\begin{enumerate}
\item Consider the matrix
\[\textbf{A} = \left[\begin{array}{cc}
    2 & 2 \\
    2 & -1
\end{array}\right]\]
Without using a computer:
\begin{enumerate}
\item Find the eigenvalues and normalized eigenvectors of \textbf{A}.
\begin{equation*}
(\textbf{A}-\lambda \textbf{I})x = 0
\end{equation*}
\item Write the spectral decomposition of \textbf{A}.
\item Verify that the determinant of \textbf{A} equals the product of its eigenvalues.
\item The trace of a square matrix equals the sum of its diagonal elements. Verify that the
trace of \textbf{A} equals the sum of its eigenvalues.
\item Is \textbf{A} orthogonal? Why or why not?
\item Is \textbf{A} positive definite? Why or why not?
\item Find \textbf{A}$^{-1}$ and determine its eigenvalues and normalized eigenvectors.
\end{enumerate}
\item Consider the matrices
These matrices are identical except for a small difference in the (2, 2) position. Also, the columns of A and B are nearly linearly dependent. Show that A-1 $\approx$ (-3)B-1. So, small changes - perhaps due to rounding - can result in substantially different inverses.
\end{enumerate}
\end{document}
